#  chapter 5. Dimensionality reduction techniques 
# Step 1.
##The data set has information obtained by the United Nations Development Program (UNDP) on various relevant topics of Human development such as Education, labour, political participation, mortality and the like. The data set contains 8 variables and 155 observations

```{r} 
human <- read.csv("~/GitHub/IODS-project/Working directory/DATA/human_.csv", header = TRUE, sep= ",")
str(human)
dim(human)
```

#Step 3

```{r}
library(corrplot)
library(GGally)
library("ggplot2")
```

```{r echo = FALSE}
renderPlot({
 d <- get(input$data)
 plot(d)
}) 
```

```{r}
pairs(human)
cor_mat <- cor(human)%>%round(digits=2)
corrplot(cor_mat, method="circle", type="upper", cl.pos="b", tl.pos="d", tl.cex = 0.6, title = "Correlation plot")
```
### According to the correlation plot, the more blue, the more correlated the variables are and the more red the more uncorrelated the variables are. Let´s say for example that life expectancy at birth extremely uncorrelated to Maternity Mortality Ratio.

# Step 3. principal component analysis (PCA)

```{r}
pca_human <- prcomp(human)
summary(pca_human)
biplot(pca_human, choices = 1:2, cex = c(0.8,1), col = c("blue", "red"))
```

# Steps 4 and 5.Standardize the variables in the human data and repeat the above analysis

```{r}
humstd <- scale(human) 
pca_humstd <- prcomp(humstd)
summary(pca_humstd)
biplot(pca_humstd, choices = 1:2, cex = c(0.8,1), col = c("blue", "red"))
```

### It seems that the wider the distance of the variables (the degree of the angle), the more uncorrelated they are. For example, maternity mortality ratio and adolescent birth are very correlated whilst expected years of life at birth and labour ratio have less correlation

# Step 6
```{r}
library(FactoMineR)
data(tea) 
str(tea)
dim(tea)

keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")
tea_time <- select(tea, one_of(keep_columns))
summary(tea_time)
str(tea_time)
```

# visualize the dataset
```{r}
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free")  + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))

mca <- MCA(tea_time, graph = FALSE)
summary(mca)
plot(mca, invisible=c("ind"), habillage = "quali")
```
